---
title: "Probit Regression with Simulated Data"
output:
  pdf_document:
    number_sections: true
geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(coda)
library(ggplot2)
library(pROC)
source("src/rwmetrop.R")
source("src/gibbs-probit.R")
```

Draw simulated dataset.
```{r}
set.seed(12345)
n <- 1000
x <- runif(n, -2, 2)
X <- cbind(1, x)
d <- ncol(X)
Beta.true <- c(-1, 1)
p.true <- pnorm(X %*% Beta.true)
y <- rbinom(n = n, size = 1, prob = p.true)
```

# Metropolis-Hastings Sampler
Prepare for Metropolis-Hastings.
``` {r}
logpost <- function(parm, Data) {
	n <- nrow(Data$X)
	d <- ncol(Data$X)
	Beta <- parm[1:d]
	Beta.prior <- sum(dnorm(Beta, 0, sqrt(hyper$var.Beta), log = TRUE))
	prob <- pnorm(Data$X %*% Beta)
	ll <- dbinom(Data$y, 1, prob, log = TRUE)
	sum(ll) + Beta.prior
}

par.init <- rep(0, d)
Data <- list(y = y, X = X)
hyper <- list(var.Beta = 1000)
```

Here the proposal variance is too small. The sampler is slow to move the chain to the posterior.
``` {r, fig.height=7}
proposal <- list(var = diag(d), scale = 0.001)
metrop.out <- rwmetrop(par.init, logpost, Data, proposal, R = 10000,
	burn = 0, thin = 1)
print(metrop.out$accept)

Beta.mcmc <- mcmc(metrop.out$par)
plot(Beta.mcmc)
acf(Beta.mcmc)
```

Here the proposal variance is too large. Draws are accepted too rarely.
``` {r, fig.height=7}
proposal <- list(var = diag(d), scale = 0.5)
metrop.out <- rwmetrop(par.init, logpost, Data, proposal, R = 10000,
	burn = 0, thin = 1)
print(metrop.out$accept)

Beta.mcmc <- mcmc(metrop.out$par)
plot(Beta.mcmc)
acf(Beta.mcmc)
```

Changing the shape of the variance seems to permit larger jumps which can still be accepted.
``` {r, fig.height=7}
proposal <- list(var = solve(t(X) %*% X), scale = 5.0)
metrop.out <- rwmetrop(par.init, logpost, Data, proposal, R = 10000,
	burn = 0, thin = 1)
print(metrop.out$accept)

Beta.mcmc <- mcmc(metrop.out$par)
plot(Beta.mcmc)
acf(Beta.mcmc)
```

Let's do some thinning to reduce the autocorrelation, and discard the first 1000 draws.
``` {r, fig.height=7}
proposal <- list(var = solve(t(X) %*% X), scale = 5.0)
metrop.out <- rwmetrop(par.init, logpost, Data, proposal, R = 10000,
	burn = 1000, thin = 10)
print(metrop.out$accept)

Beta.mcmc <- mcmc(metrop.out$par)
plot(Beta.mcmc)
acf(Beta.mcmc)
```

Compute summary of the posterior.
``` {r}
summary(Beta.mcmc)
```

Plot the saved draws of `Beta`.
``` {r}
dat.sim <- data.frame(Beta1 = as.numeric(Beta.mcmc[,1]), Beta2 = as.numeric(Beta.mcmc[,2]))
dat.true <- data.frame(Beta1 = Beta.true[1], Beta2 = Beta.true[2])
ggplot(dat.sim, aes(x=Beta1, y=Beta2)) +
	stat_density_2d(geom = "raster", aes(fill = ..density..), contour = FALSE) +
	scale_fill_gradient(low="lightcyan", high="purple") +
	ggtitle("Empirical Density of MCMC Draws") +
	geom_point(data = dat.true, aes(x = Beta1, y = Beta2))
```

Plot pointwise credible intervals for `pnorm(x^T Beta)` using the saved draws of `Beta`.
``` {r}
x.levels <- seq(-2, 2, 0.1)
dat.plot <- data.frame(
	x = x.levels,
	lo = numeric(length(x.levels)),
	mean = numeric(length(x.levels)),
	hi = numeric(length(x.levels))
)

for (idx in 1:length(x.levels)) {
	x <- x.levels[idx]
	p.draws <- pnorm(cbind(1, x) %*% t(Beta.mcmc))
	dat.plot$lo[idx] <- quantile(p.draws, probs = 0.025)
	dat.plot$mean[idx] <- mean(p.draws)
	dat.plot$hi[idx] <- quantile(p.draws, probs = 0.975)
}

ggplot(dat.plot) + 
	geom_line(aes(x = x, y = lo), lty = 2) +
	geom_line(aes(x = x, y = mean)) +
	geom_line(aes(x = x, y = hi), lty = 2) +
	ylab('p(x)')
```

Plot means from the posterior predictive distribution using the saved draws of `Beta`.
``` {r}
R.keep <- NROW(Beta.mcmc)
x.levels <- seq(-2, 2, 0.1)
dat.plot <- data.frame(
	x = x.levels,
	mean = numeric(length(x.levels))
)

for (idx in 1:length(x.levels)) {
	x <- x.levels[idx]
	p.draws <- pnorm(cbind(1, x) %*% t(Beta.mcmc))
	y.draws <- rbinom(R.keep, size = 1, prob = p.draws)
	dat.plot$mean[idx] <- mean(y.draws)
}

ggplot(dat.plot) + 
	geom_line(aes(x = x, y = mean)) +
	ylab('p(x)')
```

Use means from the posterior predictive distribution to compute ROC curve.
``` {r}
R.keep <- NROW(Beta.mcmc)
y.hat <- numeric(n)

for (idx in 1:n) {
	x <- X[idx,]
	p.draws <- pnorm(cbind(1, x) %*% t(Beta.mcmc))
	y.draws <- rbinom(R.keep, size = 1, prob = p.draws)
	y.hat[idx] <- mean(y.draws)
}

roc.out <- roc(y, y.hat)
plot(roc.out)
print(roc.out)
```

# Gibbs Sampler
Try the Gibbs sampler and check the diagnostics.
``` {r, fig.height=7}
hyper <- list(V.prior = 1000 * diag(d))
gibbs.out <- gibbs.probit(y, X, R = 1000, burn = 0, thin = 1, hyper = hyper)

Beta.mcmc <- mcmc(gibbs.out$Beta.hist)
plot(Beta.mcmc)
acf(Beta.mcmc)
```

Try a larger sample with some burning and thinning.
``` {r, fig.height=7}
hyper <- list(V.prior = 1000 * diag(d))
gibbs.out <- gibbs.probit(y, X, R = 10000, burn = 1000, thin = 10, hyper = hyper)

Beta.mcmc <- mcmc(gibbs.out$Beta.hist)
plot(Beta.mcmc)
acf(Beta.mcmc)
```

Compute summary of the posterior draws for `Beta`.
``` {r}
summary(Beta.mcmc)
```

Plot the saved draws of `Beta`.
``` {r}
dat.sim <- data.frame(Beta1 = as.numeric(Beta.mcmc[,1]), Beta2 = as.numeric(Beta.mcmc[,2]))
dat.true <- data.frame(Beta1 = Beta.true[1], Beta2 = Beta.true[2])
ggplot(dat.sim, aes(x=Beta1, y=Beta2)) +
	stat_density_2d(geom = "raster", aes(fill = ..density..), contour = FALSE) +
	scale_fill_gradient(low="lightcyan", high="purple") +
	ggtitle("Empirical Density of MCMC Draws") +
	geom_point(data = dat.true, aes(x = Beta1, y = Beta2))
```

Plot pointwise credible intervals for `pnorm(x^T Beta)` using the saved draws of `Beta`.
``` {r}
x.levels <- seq(-2, 2, 0.1)
dat.plot <- data.frame(
	x = x.levels,
	lo = numeric(length(x.levels)),
	mean = numeric(length(x.levels)),
	hi = numeric(length(x.levels))
)

for (idx in 1:length(x.levels)) {
	x <- x.levels[idx]
	p.draws <- pnorm(cbind(1, x) %*% t(Beta.mcmc))
	dat.plot$lo[idx] <- quantile(p.draws, probs = 0.025)
	dat.plot$mean[idx] <- mean(p.draws)
	dat.plot$hi[idx] <- quantile(p.draws, probs = 0.975)
}

ggplot(dat.plot) + 
	geom_line(aes(x = x, y = lo), lty = 2) +
	geom_line(aes(x = x, y = mean)) +
	geom_line(aes(x = x, y = hi), lty = 2) +
	ylab('p(x)')
```

Plot means from the posterior predictive distribution using the saved draws of `Beta`.
``` {r}
R.keep <- NROW(Beta.mcmc)
x.levels <- seq(-2, 2, 0.1)
dat.plot <- data.frame(
	x = x.levels,
	mean = numeric(length(x.levels))
)

for (idx in 1:length(x.levels)) {
	x <- x.levels[idx]
	p.draws <- pnorm(cbind(1, x) %*% t(Beta.mcmc))
	y.draws <- rbinom(R.keep, size = 1, prob = p.draws)
	dat.plot$mean[idx] <- mean(y.draws)
}

ggplot(dat.plot) + 
	geom_line(aes(x = x, y = mean)) +
	ylab('p(x)')
```

Use means from the posterior predictive distribution to compute ROC curve.
``` {r}
R.keep <- NROW(Beta.mcmc)
y.hat <- numeric(n)

for (idx in 1:n) {
	x <- X[idx,]
	p.draws <- pnorm(cbind(1, x) %*% t(Beta.mcmc))
	y.draws <- rbinom(R.keep, size = 1, prob = p.draws)
	y.hat[idx] <- mean(y.draws)
}

roc.out <- roc(y, y.hat)
plot(roc.out)
print(roc.out)
```
