---
title: "OPG Algorithm with Rcpp"
output:
  html_document:
    number_sections: true
geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(Rcpp)
library(RcppArmadillo)
library(ggplot2)
library(MASS)
```

# Read the diamonds data
``` {r}
head(diamonds)
dat <- as.data.frame(diamonds)
dat$x <- scale(dat$x)
dat$y <- scale(dat$y)
dat$z <- scale(dat$z)
dat$table <- scale(dat$table)
dat$depth <- scale(dat$depth)
dat$carat <- scale(dat$carat)
n <- nrow(diamonds)

X <- model.matrix(log(price) ~ x + y + z + depth + table + carat - 1,
	data = diamonds)
y <- log(diamonds$price)
```

# Kernel weights
Compute kernel weights between the rows of `X` and the rows of `X0`. First in R.
``` {r}
kern.weights <- function(X, X0, h)
{
	n <- NROW(X)
	p <- NCOL(X)
	n0 <- NROW(X0)
	stopifnot(p == NCOL(X0))
	W.raw <- matrix(0, n, n0)

	# Gaussian kernel
	kern <- function(u) { exp(-rowSums(u^2) / 2) }

	for (idx2 in 1:n0) {
		# Vectorized to avoid a nested loop
		U <- X - matrix(X0[idx2,], n, p, byrow = TRUE)
		W.raw[,idx2] <- kern(U / h)
	}

	W <- W.raw * matrix(1 / colSums(W.raw), n, n0, byrow = TRUE)
	return(W)
}
```

Now in Rcpp.
``` {Rcpp}
// [[Rcpp::depends(RcppArmadillo)]]
#include <RcppArmadillo.h>

// [[Rcpp::export]]
arma::vec normalize(const arma::vec& x)
{
	return x / sum(x);
}

// [[Rcpp::export]]
double gauss_kern(const arma::vec& x)
{
	return exp(-dot(x,x) / 2);
}

// [[Rcpp::export]]
arma::mat kern_weights_cpp(const arma::mat& X, const arma::mat& X0, double h)
{
	size_t n = X.n_rows;
	size_t n0 = X0.n_rows;
	size_t d = X.n_cols;
	if (d != X0.n_cols) {
		Rcpp::stop("d != X0.n_cols");
	}
	arma::mat W = arma::zeros(n, n0);

	for (size_t idx2 = 0; idx2 < n0; idx2++) {
		for (size_t idx1 = 0; idx1 < n; idx1++) {
			arma::vec u = arma::trans(X.row(idx1) - X0.row(idx2));
			W(idx1,idx2) = gauss_kern(u / h) / (h*d);
		}
		if (arma::any(W.col(idx2) > 0)) {
			W.col(idx2) = normalize(W.col(idx2));
		}
	}

	return W;
}
```

Check results to make sure they match.
``` {r}
x0 <- t(X[1,])
head( ww1 <- kern.weights(X, X0 = x0, h = 1) )
head( ww2 <- kern_weights_cpp(X, X0 = x0, h = 1) )
print( sum(abs(ww1 - ww2)) )
```

# Local coefficients
Compute local regression coefficients at point `x0`. First in R.
``` {r}
solve.local <- function(y, X, x0, h, ww)
{
	p <- length(x0)
	n <- length(y)
	ww <- as.numeric(ww)

	X0 <- matrix(x0, n, p, byrow = TRUE)
	Z <- cbind(1, X - X0)
	xi <- ginv(t(Z) %*% (ww * Z)) %*% (t(Z) %*% (ww * y))
	return(xi)
}
```

Now in Rcpp.
``` {Rcpp}
// [[Rcpp::depends(RcppArmadillo)]]
#include <RcppArmadillo.h>

// [[Rcpp::export]]
arma::vec solve_local_cpp(const arma::vec& y, const arma::mat& X,
	const arma::vec& x0, double h, const arma::vec& ww)
{
	size_t p = x0.n_elem;
	size_t n = y.n_elem;

	arma::mat X0 = arma::repmat(arma::trans(x0), n, 1);
	arma::mat Z = arma::join_rows(arma::ones(n,1), X - X0);
	arma::mat WW = repmat(ww, 1, p+1);
	arma::vec xi = arma::pinv(arma::trans(Z) * (WW % Z)) * (arma::trans(Z) * (ww % y));
	return xi;
}
```

Check results to make sure they match.
``` {r}
x0 <- t(X[1,])
ww <- kern.weights(X, X0 = x0, h = 1)

solve.local(y, X, x0, h = 1, ww)
solve_local_cpp(y, X, x0, h = 1, ww)
```

# Compare timings

``` {r}
h <- 1
idx <- 1:500

# Do both operations in R
st <- Sys.time()
for (i in idx) {
	x0 <- t(X[i,])
	ww <- kern.weights(X, X0 = x0, h = h)	
	xi <- solve.local(y, X, x0, h = h, ww)
}
cat("Elapsed secs ", as.numeric(Sys.time() - st, units = "secs"), "\n")

# Do weights in R and local coeffs in C++
st <- Sys.time()
for (i in idx) {
	x0 <- t(X[i,])
	ww <- kern.weights(X, X0 = x0, h = h)	
	xi <- solve_local_cpp(y, X, x0, h = h, ww)
}
cat("Elapsed secs ", as.numeric(Sys.time() - st, units = "secs"), "\n")

# Do weights in C++ and local coeffs in R
st <- Sys.time()
for (i in idx) {
	x0 <- t(X[i,])
	ww <- kern_weights_cpp(X, X0 = x0, h = h)	
	xi <- solve.local(y, X, x0, h = h, ww)
}
cat("Elapsed secs ", as.numeric(Sys.time() - st, units = "secs"), "\n")

# Do both operations in C++
st <- Sys.time()
for (i in idx) {
	x0 <- t(X[i,])
	ww <- kern_weights_cpp(X, X0 = x0, h = h)	
	xi <- solve_local_cpp(y, X, x0, h = h, ww)
}
cat("Elapsed secs ", as.numeric(Sys.time() - st, units = "secs"), "\n")
```

A complete OPG code is available in the workshop materials.
