---
title: "Practical Rcpp"
output:
  html_document:
    df_print: paged
  pdf_document:
    number_sections: yes
geometry: left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(Rcpp)
library(RcppArmadillo)
```

# Modular Code Practice
A good practice for Rcpp programmers is to initially write code in R, ensure that it is correct, and then write a C++ version. The C++ version can then be easily compared to the R version.

Suppose $X_1, \ldots, X_n \stackrel{\text{iid}}{\sim} \text{Categ}(1, \ldots, K; \pi_1, \ldots, \pi_K)$. Let us compute the density in both R and Rcpp.

``` {r}
dcateg <- function(x, Pi, log = FALSE)
{
	if (log) {
		log(Pi[x])
	} else {
		Pi[x]
	}
}
```

``` {Rcpp}
// [[Rcpp::depends("RcppArmadillo")]]
#include <RcppArmadillo.h>

// [[Rcpp::export]]
arma::vec dcateg_cpp(const arma::uvec& x, const arma::vec& Pi, bool log = false)
{
	if (log) {
		return arma::log(Pi(x-1));
	} else {
		return Pi(x-1);
	}
}
```

``` {r}
set.seed(1234)
J <- 5
Pi <- 1:J / sum(1:J)
x <- sample(1:J, size = 20, replace = TRUE, prob = Pi)
```

``` {r}
dcateg(x, Pi, log = FALSE)
dcateg_cpp(x, Pi, log = FALSE)

dcateg(x, Pi, log = TRUE)
dcateg_cpp(x, Pi, log = TRUE)
```

When functions use the random number generator, comparing R and C++ versions is much easier when the same numbers can be generated in both settings.

``` {Rcpp}
#include <Rcpp.h>

// [[Rcpp::export]]
Rcpp::NumericVector draw(unsigned int n, double mu, double sigma)
{
	return Rcpp::rnorm(n, mu, sigma);
}
```

``` {r}
set.seed(1234)
rnorm(20, 0, 3)

set.seed(1234)
draw(20, 0, 3)
```

Now consider a small simulation to study the sampling mean $\bar{X}$ from our categorical sample.
``` {r}
sim <- function(R, n, Pi)
{
	res <- numeric(R)
	for (r in 1:R)	 {
		x <- sample(1:J, size = n, replace = TRUE, prob = Pi)
		res[r] <- mean(x)
	}
	return(res)
}

xbar <- sim(R = 10000, n = 20, Pi)
quantile(xbar, probs = seq(0, 1, 0.1))
```

There appears not to be an Rcpp version of `sample`. To work around this, we will write our own drawing function in Rcpp.

``` {Rcpp}
// [[Rcpp::depends("RcppArmadillo")]]
#include <RcppArmadillo.h>

// Return the index of the draw, 0-based
// Probabilties must be normalized for this method to work
long unsigned int rcateg(const arma::vec& Pi)
{
	size_t J = Pi.size();
	double u = R::runif(0, 1);
	double qq = 0;

	for (size_t j = 0; j < J; j++) {
		qq += Pi(j);
		if (u < qq) {
			return j;
		}
	}

	return J-1;
}

// [[Rcpp::export]]
arma::uvec rcateg(long unsigned int n, const arma::vec& Pi)
{
	arma::uvec x(n);
	for (size_t i = 0; i < n; i++) {
		x(i) = rcateg(Pi);
	}
	return x + 1;
}

// [[Rcpp::export]]
arma::vec sim_cpp(unsigned int R, unsigned int n, const arma::vec& Pi)
{
	arma::vec res(R);
	for (unsigned int r = 0; r < R; r++) {
		arma::uvec x = rcateg(n, Pi);
		res(r) = double(sum(x)) / n;
	}
	return(res);
}
```

Redo the simulation in R, drawing from `rcateg` instead of `sample`.
``` {r}
sim2 <- function(R, n, Pi)
{
	res <- numeric(R)
	for (r in 1:R) {
		x <- rcateg(n, Pi)
		res[r] <- mean(x)
	}
	return(res)
}
```

``` {r}
set.seed(1234)
sim2(R = 100, n = 20, Pi)

set.seed(1234)
as.numeric(sim_cpp(R = 100, n = 20, Pi))
```

In a complicated R program, it is helpful to express major operations as functions. These functions can be converted to Rcpp as needed. We will see an example outside of this document.

# Debugging with Print Statements
Print statements are a simple way to debug. Consider computing the log-likelihood of $X_1, \ldots, X_n \stackrel{\text{iid}}{\sim} \text{N}(\mu, \sigma^2)$.

``` {Rcpp}
#include <Rcpp.h>

// [[Rcpp::export]]
double loglik(const Rcpp::NumericVector& x, double mu, double sigma)
{
	double ll = 0;
	for (unsigned int i = 0; i < x.size(); i++) {
		ll += R::dnorm(x(i), mu, sigma, true);
	}
	
	return ll;
}
```

Underflow can occur if we work on the probability scale and take the log at the end. Printing some of the intermediate steps helps to see this.
``` {Rcpp}
#include <Rcpp.h>

// [[Rcpp::export]]
double likelihood(const Rcpp::NumericVector& x, double mu, double sigma)
{
	double lik = 1;
	for (unsigned int i = 0; i < x.size(); i++) {
		lik *= R::dnorm(x(i), mu, sigma, false);
		if (lik < 1e-320) {
			Rprintf("At step %d, lik = %g\\n", i, lik);
		}
	}
	
	return lik;
}
```

``` {r}
set.seed(1234)

n <- 50
x <- rnorm(n, mean = 2, sd = 1)
log(likelihood(x, mu = 2, sigma = 1))
loglik(x, mu = 2, sigma = 1)

n <- 530
x <- rnorm(n, mean = 2, sd = 1)
log(likelihood(x, mu = 2, sigma = 1))
loglik(x, mu = 2, sigma = 1)
```

# Exceptions
Can throw exceptions from C++ up to R.
``` {Rcpp}
#include <Rcpp.h>

double my_sqrt(double x) {
	if (x < 0) {
		Rcpp::stop("x was negative!");
	}
	return pow(x, 0.5);
}

// [[Rcpp::export]]
Rcpp::NumericVector my_sqrt(const Rcpp::NumericVector& x)
{
	Rcpp::NumericVector z(x.size());
	for (unsigned int i = 0; i < x.size(); i++) {
		z(i) = my_sqrt(x(i));
	}
	return z;
}
```

``` {r, error=TRUE}
set.seed(1234)
x <- rnorm(10, mean = 0, sd = 1)
my_sqrt(x)
```

Catch an exception in C++ and handle it there.
``` {Rcpp}
#include <Rcpp.h>

double my_sqrt(double x) {
	if (x < 0) {
		Rcpp::stop("x was negative!");
	}
	return pow(x, 0.5);
}

// [[Rcpp::export]]
Rcpp::NumericVector my_sqrt2(const Rcpp::NumericVector& x)
{
	Rcpp::NumericVector z(x.size());
	z.fill(NAN);
	for (unsigned int i = 0; i < x.size(); i++) {
		try {
			z(i) = my_sqrt(x(i));
		} catch (const std::exception& e) {
			Rprintf("%s\\n", e.what());
		}
	}
	return z;
}
```

``` {r}
round(x, 3)
my_sqrt2(x)
x^(0.5)
```

Catch an exception in R which was thrown from C++.
``` {r}
result <- tryCatch({
	my_sqrt(x)
}, error = function(e) {
	cat("Caught an error:\n")
	print(e)

	# Take the sqrt of abs(x) instead
	my_sqrt(abs(x))
})
```

# Logging Progress
It can be helpful to log important steps with a timestamp.
``` {r}
logger <- function(msg, ...) {
	sys.time <- as.character(Sys.time())
	cat(sys.time, "-", sprintf(msg, ...))
}
```

``` {r}
logger("Starting MCMC\n")
Sys.sleep(1)
logger("Finished MCMC\n")
```

We can also log important steps in C++ code.
``` {Rcpp}
#include <Rcpp.h>
#include <unistd.h>

Rcpp::Datetime timestamp() {
  time_t t;
  time(&t);
  return Rcpp::Datetime(t);
}

Rcpp::String timestring() {
	char buffer[50];
	const Rcpp::Datetime& dt = timestamp();
	sprintf(buffer, "%04d-%02d-%02d %02d:%02d:%02d", dt.getYear(), dt.getMonth(),
		dt.getDay(), dt.getHours(), dt.getMinutes(), dt.getSeconds());
	return Rcpp::String(buffer);
}

// [[Rcpp::export]]
void test(unsigned int elapsed) {
	Rprintf("%s - Starting MCMC\\n", timestring().get_cstring());
	sleep(elapsed);
	Rprintf("%s - Finished MCMC\\n", timestring().get_cstring());
}
```

``` {r}
test(5)
```

# Measuring Speed
We have already seen logging, which can be used to measure elapsed time. We can also wrap function calls with `system.time` (and related functions).

``` {r}
system.time(sim2(R = 10000, n = 20, Pi))
system.time(sim_cpp(R = 10000, n = 20, Pi))
```

Or use `Sys.time` to take measurements at two points in the code, and compute the difference.
``` {r}
st1 <- Sys.time()
res1 <- sim2(R = 10000, n = 20, Pi)
end1 <- Sys.time()

st2 <- Sys.time()
res2 <- sim_cpp(R = 10000, n = 20, Pi)
end2 <- Sys.time()

# Print difftime objects
print(end1 - st1)
print(end2 - st2)

# Convert to raw numbers which are convenient for other calculations
as.numeric(end1 - st1, units = "secs")
as.numeric(end2 - st2, units = "secs")
```

Microbenchmarking is convenient for repeating the same calculation many times and extracting statistics on run times.

``` {r}
library(microbenchmark)

mb <- microbenchmark(
	sim2(R = 10000, n = 20, Pi),
	sim_cpp(R = 10000, n = 20, Pi),
	times = 100)
print(mb)
```

# Measuring Memory Use
The `pryr` package provides some functions to keep track of memory usage.

``` {r}
library(pryr)

mem_used()

X <- matrix(NA, 10000, 10000)
mem_used()

object_size(X)

mem_change(Y <- matrix(NA, 1000, 1000))
mem_change(Y <- matrix(NA, 2000, 2000))

mem_change(rm(X))
mem_change(rm(Y))
gc()

mem_used()
```
