---
title: "Toeplitz Example With Bytecode Compiler"
output:
  pdf_document:
    number_sections: true
geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(Rcpp)
library(microbenchmark)
library(compiler)
```

First, turn off automatic bytecode compilation.
``` {r}
enableJIT(0)
```

Nested for loops in R.
```{r}
toeplitz.for <- function(x) {
	p <- length(x)
	A <- matrix(NA, p, p)
	for (i in 1:p) {
		for (j in 1:p) {
			A[i,j] <- x[abs(j-i) + 1]
		}
	}
	return(A)
}
```

Vectorize one of the loops.
```{r}
toeplitz.vectfor <- function(x) {
	p <- length(x)
	A <- matrix(NA, p, p)
	for (i in 1:p) {
		A[i,] <- x[abs(1:p - i) + 1]
	}
	return(A)
}
```

Use `rbind` to append each new row to previously computed rows.
``` {r}
toeplitz.rbind <- function(x) {
	p <- length(x)
	A <- matrix(NA, 0, p)
	for (j in 1:p){
		idx <- abs(1:p - j) + 1
		A <- rbind(A, x[idx])
	}
	return(A)
}
```

Use the `outer` function to compose the matrix.
``` {r}
toeplitz.outer <- function(x) {
	p <- length(x)
	A <- matrix(NA, p, p)
	outer(X = 1:p, Y = 1:p, FUN = function(i,j) {
		x[abs(i - j) + 1]
	})
}
```

Use the R functions `row` and `col` to compose the result via matrix algebra.
``` {r}
toeplitz.matrix <- function (x) 
{
	p <- length(x)
	A <- matrix(NA, p, p)
	matrix(x[abs(col(A) - row(A)) + 1], p, p)
}
```

Nested for-loops in Rcpp.
``` {Rcpp}
#include <Rcpp.h>

// [[Rcpp::export]]
Rcpp::NumericMatrix toeplitz_cpp(const Rcpp::NumericVector& x) 
{
	int p = x.size();
	Rcpp::NumericMatrix A(p, p);
	for (int j = 0; j < p; j++) {
		for (int i = 0; i < p; i++) {
			A(i,j) = x[abs(j-i)];
		}
	}
	return A;
}
```

Test the Toeplitz functions to make sure they all return approximately the same result.
``` {r}
x <- 1:5
print(x)
toeplitz_cpp(x)

stopifnot(max(toeplitz.for(x) - toeplitz_cpp(x)) < 1e-10)
stopifnot(max(toeplitz.vectfor(x) - toeplitz_cpp(x)) < 1e-10)
stopifnot(max(toeplitz.rbind(x) - toeplitz_cpp(x)) < 1e-10)
stopifnot(max(toeplitz.outer(x) - toeplitz_cpp(x)) < 1e-10)
stopifnot(max(toeplitz.matrix(x) - toeplitz_cpp(x)) < 1e-10)
```

Run microbenchmarking to compare the performance of the functions.
``` {r}
options(width = 100)
x <- 1:1000
mb <- microbenchmark(
  toeplitz.for(x), toeplitz.vectfor(x), toeplitz.rbind(x), toeplitz.outer(x),
  toeplitz.matrix(x), toeplitz_cpp(x), times = 50
)
print(mb)
```

Try the bytecode compiler on all Toeplitz functions.
``` {r}
options(width = 100)

toeplitz.for2 <- cmpfun(toeplitz.for)
toeplitz.vectfor2 <- cmpfun(toeplitz.vectfor)
toeplitz.rbind2 <- cmpfun(toeplitz.rbind)
toeplitz.outer2 <- cmpfun(toeplitz.outer)
toeplitz.matrix2 <- cmpfun(toeplitz.matrix)
toeplitz_cpp2 <- cmpfun(toeplitz_cpp)

x <- 1:1000
mb <- microbenchmark(
  toeplitz.for2(x), toeplitz.vectfor2(x), toeplitz.rbind2(x),
  toeplitz.outer2(x), toeplitz.matrix2(x), toeplitz_cpp2(x),
  times = 50
)
print(mb)
```
