---
title: 'SLR Example: Album Sales Data'
author: "NKN"
date: "January 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load relevant libraries
```{r}
library(readr)
library(ggplot2)
library(xtable)
```

#Read the data
```{r}
AlbumSales1 <- read_csv("../../data/csv/AlbumSales1.csv")
head(AlbumSales1)
names(AlbumSales1)
summary(AlbumSales1)
library(knitr)
kable(summary(AlbumSales1))
```

#ScatterPlot
```{r}
attach(AlbumSales1)
plot(adverts,sales)
title("Scatterplot of Sales Figures vs Advertisement Expenditure")

```


#Fancy Scatterplot
```{r}
attach(AlbumSales1)
scatter <- ggplot(AlbumSales1, aes(adverts, sales))
scatter + geom_point() + geom_smooth() + labs(x = "Ad Exp", y = "Sales") + ggtitle("      Scatterplot of Sales Figures vs Advertisement Expenditure")

```


#Fancy Scatterplot with the Regression Line added
```{r}
attach(AlbumSales1)
scatter <- ggplot(AlbumSales1, aes(adverts, sales))
scatter + geom_point() + 
	geom_smooth() +
	geom_smooth(method = "lm", colour = "Red") + 
	labs(x = "Ad Exp", y = "Sales") + 
	ggtitle("      Scatterplot of Sales Figures vs Advertisement Expenditure")

```


#Fitting a Simple Linear Regression
```{r}
AlbumSales.out1 <- lm(sales ~ adverts, data = AlbumSales1)
summary(AlbumSales.out1)

```

# What are all these numbers:

$$ \mbox{Residual Standard Error} = 65.99 = \sqrt{\frac{ \mbox{ Residual Sum of Squares}}{198} } $$

$$ \mbox{Therefore,  RSS=}  198* (65.99)^2 = 198*4354.68=862226.7$$



$$ \mbox{Now, we can find SSTotal from RSS and } R^2: $$
$$0.3346=R^2=1-\frac{RSS}{SST} \implies SST= \frac{RSS}{(1-R^2)}= \frac{862226.7}{1-0.3346}=1295802;  \mbox{   and  SSModel=SST-RSS=}1295802-862227=433575  $$ 

$$ \mbox{ The F-statistics: } \frac{ \mbox{MSModel} }{ \mbox{MSError}} = \frac{ \mbox{SSModel/DF} }{ \mbox{RSS/DF}} =
\frac{ \mbox{433575/1} }{ \mbox{862227/198}} = 99.59 $$

# Compute the p-value for the above F-statistic
# For small p-values, we can find the upper bound as needed by some publications as follows.
```{r}
pvalue <- 1-pf(99.59,1,198)
formatC(pvalue, format = "e", digits = 8)
print(pvalue)
```

Now, we can construct the traditional ANOVA:

```{r xtable, results="asis"}
library(xtable)
tab <- xtable(summary(AlbumSales.out1)$coef, digits=c(0, 2, 2, 1, 2))
print(tab, type="html")
```


Check your results below using R's anova function



#Check your results below using R's anova function
```{r}
summary(AlbumSales.out1)
anova(AlbumSales.out1)
```

