---
title: 'Multiple Comparisons: Vibrations Data'
author: "NKN"
date: "Dec 5, 2018"
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read Motor.csv data

Reading data, converting Brand to a factor, and some housekeeping.

```{r}
library(readxl)
library("multcomp")
motor <- read.csv("../../data/csv/motor.csv")
motor$vib <- motor$vibration  
motor$Brand <- as.factor(motor$Brand)
```

## Compute ANOVA, Individual Confidence Intervals 

```{r}
attach(motor)
motor.aov <- aov(vib~Brand)
motor.aov.glht <- glht(motor.aov, linfct = mcp(Brand = "Tukey"))
summary(motor.aov.glht,test=univariate())
plot(motor.aov.glht)
detach(motor)
```

## Compute Simultaneous Confidence Intervals

The following lines can be used to compute the p-values
for testing the equality pairwise means using different
methods of adjusting the p-values. 


```{r}
summary(motor.aov.glht)
summary(motor.aov.glht,test=adjusted("Shaffer"))
summary(motor.aov.glht,test=adjusted("bonferroni"))
```

## Undjusted CIs for pairwise differences

The un-adjusted confidence intervals for pairwise
differences of means are obtained below.

```{r}
confint(motor.aov.glht,level=0.95,calpha = univariate_calpha())
```

## Other contrasts

We can look at the contrasts that compare different brands to a reference brand as follows:

```{r}
motor.aov.glht.Dun <- glht(motor.aov, linfct = mcp(Brand = "Dunnett"))
confint(motor.aov.glht.Dun,level=0.95,calpha = univariate_calpha())
summary(motor.aov.glht.Dun)
```

Suppose we want to compare abrands 1-4 to 5, instead of the above. Then, we need to change the reflevel. This can be done as follows. 

```{r}
motor$Brand <- relevel(motor$Brand, ref="5")
levels(motor$Brand) 
# now "5" becomes the reference level -- note that 5 is listed first
```

## Using Brand 5 as reference

Now the same code as before will compare the brands to the new reference level 5.

```{r}
motor$Brand <- relevel(motor$Brand, ref="5")
levels(motor$Brand) 
motor.aov2 <- aov(vib~Brand,data=motor)
summary(motor.aov2)
motor.aov.glht.Dun2 <- glht(motor.aov2, linfct = mcp(Brand = "Dunnett"))
confint(motor.aov.glht.Dun2,level=0.95,calpha = univariate_calpha())
summary(motor.aov.glht.Dun2)
```


