---
title: "Bootstrap Example"
output:
  pdf_document:
    number_sections: true
geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this example, we will bootstrap the distribution of coefficient estimates in a linear regression. We can compare the bootstrapped distribution to the theoretical distribution (which is well known).

```{r}
library(ggplot2)
library(mvtnorm)
```

Generate some data from ordinary least squares model.
```{r}
set.seed(1234)
n <- 400
x <- runif(n, -2, 2)
X <- cbind(1, x)
Beta.true <- c(-1, 1)
sigma.true <- 0.5
y <- rnorm(n, X %*% Beta.true, sigma.true)
```

Run the bootstrap.
```{r}
set.seed(1234)
lm.out <- lm(y ~ x)
Beta.hat <- coef(lm.out)
V.hat <- vcov(lm.out)

B <- 500
Beta.boot <- matrix(NA, B, 2)

for (b in 1:B) {
	idx.boot <- sample(1:n, size = n, replace = TRUE)
	lm.boot <- lm(y ~ x, subset = idx.boot)
	Beta.boot[b,] <- coef(lm.boot)
}
```

Compare the distribution of `Beta.boot` from bootstrap with the theoretical distribution. Center the theoretical distribution at `Beta.hat` rather than `Beta.true`, to see if we got the variability correct (acknowledging that estimates themselves will differ from the truth).
``` {r, fig.show='hold', fig.width=4.25}
V.true <- sigma.true^2 * solve(t(X) %*% X)

hist(Beta.boot[,1], freq = FALSE)
curve(dnorm(x, Beta.hat[1], sqrt(V.true[1,1])), add = TRUE)

hist(Beta.boot[,2], freq = FALSE)
curve(dnorm(x, Beta.hat[2], sqrt(V.true[2,2])), add = TRUE)
```


Plot of 2-d empirical density from simulation versus the normal distribution from large sample theory.
``` {r, fig.show='hold', fig.width=4.25, fig.height=3.25}
dat.sim <- data.frame(x = Beta.boot[,1], y = Beta.boot[,2])

dat.mvn <- expand.grid(
	x = seq(min(dat.sim$x), max(dat.sim$x), length.out = 100),
	y = seq(min(dat.sim$y), max(dat.sim$y), length.out = 100))
dat.mvn$dens <- dmvnorm(dat.mvn, mean = Beta.hat, sigma = V.true)

ggplot(dat.sim, aes(x=x, y=y)) + 
	stat_density_2d(geom = "raster", aes(fill = ..density..), contour = FALSE) +
	scale_fill_gradient(low="lightcyan", high="purple") +
	ggtitle("Empirical Density of MLE")

ggplot(dat.mvn, aes(x=x, y=y)) + 
    geom_raster(aes(fill = dens)) +
	scale_fill_gradient(low="lightcyan", high="purple") +
	ggtitle("MVN Density Based on MLE")
```
