---
title: "Cross-validation Example"
output:
  pdf_document:
    number_sections: true
geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generate some data. If an observation has Euclidean distance from the origin between sqrt(1/4) and 1, then its in class 1; otherwise it's in class 0.
```{r}
library(ggplot2)
library(gridExtra)

set.seed(1234)
n <- 20^2
x1 <- seq(-sqrt(2), sqrt(2), length.out = sqrt(n))
x2 <- seq(-sqrt(2), sqrt(2), length.out = sqrt(n))
X <- as.matrix(expand.grid(x1, x2))
r2 <- X[,1]^2 + X[,2]^2
y <- as.integer(r2 < 1 & r2 > 1/4)
dat <- data.frame(x1 = X[,1], x2 = X[,2], y = y)

color <- numeric(n)
color[dat$y == 1] <- "black"
color[dat$y == 0] <- "red"

ggplot(dat, aes(x1, x2)) +
    geom_point(color = color) +
	ggtitle("Observed Classes")+
	annotate("path",
		x = cos(seq(0, 2*pi, length.out=100)),
		y = sin(seq(0, 2*pi, length.out=100))) +
	annotate("path",
		x = sqrt(1/4) * cos(seq(0, 2*pi, length.out=100)),
		y = sqrt(1/4) * sin(seq(0, 2*pi, length.out=100)))
```

Write a simple N-nearest neighbors algorithm.
``` {r}
nearest.neighbors <- function(y, X, X.new, N)
{
	stopifnot(ncol(X) == 2)
	stopifnot(ncol(X.new) == 2)
	n <- nrow(X)
	n.new <- nrow(X.new)
	y.hat <- integer(n.new)

	for (i in 1:n.new) {
		Z <- X - matrix(X.new[i,], n, 2, byrow = TRUE)
		D <- sqrt(rowSums(Z^2))
		idx.ordered <- order(D)
		y.hat[i] <- mean(y[idx.ordered[1:N]]) > 0.5
	}

	return(y.hat)
}
```

Draw another dataset which is not necessarily on the original grid.
``` {r}
n.new <- 810
X.new <- cbind(
	runif(n.new, -sqrt(2), sqrt(2)),
	runif(n.new, -sqrt(2), sqrt(2))
)
```

`N = 1` might be a bit too small.
``` {R}
y.hat <- nearest.neighbors(dat$y, X, X.new, N = 1)

color <- numeric(n.new)
color[y.hat == 1] <- "black"
color[y.hat == 0] <- "red"

dat.new <- data.frame(x1 = X.new[,1], x2 = X.new[,2], y.hat = y.hat)
ggplot(dat.new, aes(x1, x2)) +
    geom_point(color = color) +
	ggtitle("Predicted Classes") +
	annotate("path",
		x = cos(seq(0, 2*pi, length.out=100)),
		y = sin(seq(0, 2*pi, length.out=100))) +
	annotate("path",
		x = sqrt(1/4) * cos(seq(0, 2*pi, length.out=100)),
		y = sqrt(1/4) * sin(seq(0, 2*pi, length.out=100)))
```

`N = 70` seems to be a poor choice.
``` {R}
y.hat <- nearest.neighbors(dat$y, X, X.new, N = 70)

color <- numeric(n.new)
color[y.hat == 1] <- "black"
color[y.hat == 0] <- "red"

dat.new <- data.frame(x1 = X.new[,1], x2 = X.new[,2], y.hat = y.hat)
ggplot(dat.new, aes(x1, x2)) +
    geom_point(color = color) +
	ggtitle("Predicted Classes") +
	annotate("path",
		x = cos(seq(0, 2*pi, length.out=100)),
		y = sin(seq(0, 2*pi, length.out=100))) +
	annotate("path",
		x = sqrt(1/4) * cos(seq(0, 2*pi, length.out=100)),
		y = sqrt(1/4) * sin(seq(0, 2*pi, length.out=100)))
```

Run cross-validation.
``` {r}
set.seed(1234)
idx.train <- sample(1:n, size = ceiling(0.8*n))
idx.test <- setdiff(1:n, idx.train)
color <- ifelse(1:n %in% idx.train, "blue", "darkorange")

# Set up folds for cross-validation
K <- 10
fold.grp <- sample((1:n - 1) %% K + 1)
folds <- split(1:n, f = fold.grp)

N.levels <- 1:70
na <- rep(NA, length(N.levels))
tab <- data.frame(N = N.levels, sape = na)

for (idx1 in 1:length(N.levels)) {
	N <- N.levels[idx1]
	y.hat.cv <- rep(NA, n)
	for (k in 1:K) {
		idx.fold <- folds[[k]]
		idx.notfold <- as.integer(unlist(folds[-k]))
		y.hat.cv[idx.fold] <- nearest.neighbors(dat$y[idx.notfold], X[idx.notfold,],
			X[idx.fold,], N = N)
	}
	tab$sape[idx1] <- sum(abs(y - y.hat.cv))
}

ggplot(tab, aes(N, sape)) + geom_line(lwd = 1.1)

N.cv <- N.levels[which.min(tab$sape)]
cat("Cross-validation selected degree", N.cv, "\n")
```

Results for out choice of N from cross-validation
``` {R}
y.hat <- nearest.neighbors(dat$y, X, X.new, N = N.cv)

color <- numeric(n.new)
color[y.hat == 1] <- "black"
color[y.hat == 0] <- "red"

dat.new <- data.frame(x1 = X.new[,1], x2 = X.new[,2], y.hat = y.hat)
ggplot(dat.new, aes(x1, x2)) +
    geom_point(color = color) +
	ggtitle("Predicted Classes") +
	annotate("path",
		x = cos(seq(0, 2*pi, length.out=100)),
		y = sin(seq(0, 2*pi, length.out=100))) +
	annotate("path",
		x = sqrt(1/4) * cos(seq(0, 2*pi, length.out=100)),
		y = sqrt(1/4) * sin(seq(0, 2*pi, length.out=100)))
```

