---
title: "Data Manipulation with dplyr"
output:
  pdf_document:
    number_sections: true
geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(dplyr)
```

# Working with a Single Dataset
``` {r}
data(mpg, package = "ggplot2")
tail(mpg)
```

## Filter Rows
Use `filter` to choose rows from a dataset.
``` {r}
mpg %>%
	filter(manufacturer == "toyota")

mpg %>%
	filter(manufacturer == "toyota") %>%
	filter(class == "compact")

mpg %>%
	filter(manufacturer == "toyota" & model == "corolla")
```

## Select Columns
Use `select` to choose columns from a dataset.
``` {r}
mpg %>% 
	filter(manufacturer == "toyota" & model == "corolla") %>%
	select(year, trans, cty, hwy)

mpg %>% 
	filter(manufacturer == "toyota" & model == "corolla") %>%
	select(manufacturer:cyl)

mpg %>%
	filter(manufacturer == "toyota" & model == "corolla") %>%
	select(-(manufacturer:cyl))
```

## Mutate Columns
Use `mutate` to transform variables (columns) in a dataset. Transformations can be functions of more than one column.
``` {r}
mpg %>%
	filter(manufacturer == "toyota" & model == "corolla") %>%
	mutate(cty_kpg = 1.6 * cty, hwy_kpg = 1.6 * hwy) %>%
	mutate(avg_mpg = 0.5 * cty + 0.5 * hwy) %>%
	select(-manufacturer, -model, -trans)
```

## Arrange (Sort) Rows
Use `arrange` to sort a dataset.
``` {r}
mpg %>% 
	filter(manufacturer == "toyota" & class == "compact") %>%
	arrange(year) %>%
	select(manufacturer:cyl)

mpg %>% 
	filter(manufacturer == "toyota" & class == "compact") %>%
	arrange(desc(year)) %>%
	select(manufacturer:cyl)

mpg %>% 
	filter(manufacturer == "toyota" & class == "compact") %>%
	arrange(desc(year), desc(cyl), desc(displ)) %>%
	select(manufacturer:cyl)
```

## Group By
Use `group_by` and `summarize` to group rows together and compute summaries, such as the sum or average of a variable.
``` {r}
mpg %>%
	group_by(manufacturer) %>%
	summarize(count = n())

mpg %>%
	group_by(class)  %>%
	summarize(count = n(), avg_cty = mean(cty), sd_cty = sd(cty)) %>%
	arrange(desc(avg_cty))


```

# Working with Relational Data
The `nycflights13` dataset is featured in chapter 13 of the book *R for Data Science*.
``` {r}
library(nycflights13)

head(airlines)
head(airports)
head(flights)
head(planes)
head(weather)
```

## Inner Joins
Let's do an `inner_join` between flight destinations and airports
``` {r}
flights %>% 
	inner_join(airports, by = c("dest" = "faa")) %>%
	select(year, month, day, dep_time, name)

flights %>% 
	inner_join(airports, by = c("dest" = "faa")) %>%
	select(year, month, day, dep_time, name) %>%
	group_by(name) %>%
	summarize(count = n())
```

## Outer Joins
To demonstrate outer joins, let's set up a table with prices of several models of cars.

* An `inner_join` only returns rows that have a match in both datasets.
* A `left_join` is one type of outer join; it returns all rows in the "left" dataset and assumes `NA` for columns from the "right" dataset which did not have a match.
* On the other hand, a `right_join` returns all rows in the "right" dataset...
``` {r}
options(width=100)
prices <- tibble(
	mfr = c("toyota", "toyota", "toyota"),
	model = c("corolla", "camry", "prius"),
	price = c(18000, 22000, 23000)
)

mpg %>%
	inner_join(prices, by = c("manufacturer" = "mfr", "model" = "model")) %>%
	select(-trans, -class)

mpg %>%
	left_join(prices, by = c("manufacturer" = "mfr", "model" = "model")) %>%
	select(-trans, -class)

mpg %>%
	left_join(prices, by = c("manufacturer" = "mfr", "model" = "model")) %>%
	select(-trans, -class) %>%
	filter(manufacturer == "toyota")

mpg %>%
	right_join(prices, by = c("manufacturer" = "mfr", "model" = "model")) %>%
	select(-trans, -class)
```

## Cartesian Joins
A cartesian join simply joins every row of the left dataset to every row of the right dataset. Let's create two small tables to demonstrate a cartesian join. Try `merge` in base R and `expand` in the tidyverse.
``` {r}
library(tidyr)

dat1 <- data.frame(my_factor = factor(c('a','b','c','d')))
dat2 <- data.frame(my_int = 1:3)
merge(dat1, dat2)

dat1 %>% 
	expand(my_factor, my_int = dat2$my_int)
```
