---
title: "Numerical Tools"
output:
  pdf_document:
    number_sections: true
geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(ggplot2)
```

# Optimization
``` {r}
f <- function(x) { cos(0.5*pi*x[1]) + cos(0.5*pi*x[2]) }

dat.plot <- expand.grid(
    x = seq(-2, 2, 0.05),
    y = seq(-2, 2, 0.05))
dat.plot$f <- apply(dat.plot, 1, f)

ggplot(dat.plot, aes(x=x, y=y)) + 
	geom_raster(aes(fill = f)) +
	geom_contour(aes(z = f), binwidth = 0.2) +
	scale_fill_gradient(low="white", high="blue")

optim.out <- optim(c(-1,1), f, method = "BFGS", control = list(fnscale = -1))
print(optim.out)
```

# Integration
If $X \sim \text{Beta}(a,b)$, then moment generating function is
\begin{align*}
M_X(t) = \int_0^1 e^{tx} \frac{x^{a-1} (1-x)^{b-1}}{ B(a,b) } dx.
\end{align*}

To evaluate this numerically at $a = 2, b = 1.5, t = 1$,
``` {r}
f <- function(x, a, b, t) {
    exp(t*x) * x^(a - 1) * (1-x)^(b-1) / beta(a,b)
}
```

A more stable version of the function
``` {r}
g <- function(x, a, b, t) {
    exp( t*x + (a - 1)*log(x) + (b-1)*log(1-x) - lbeta(a,b) )
}
```

``` {r}
integrate(f, lower = 0, upper = 1, a = 2, b = 1.5, t = 1)
integrate(g, lower = 0, upper = 1, a = 2, b = 1.5, t = 1)
```

# Root Finding
Let $f$ be the density of the F-distribution with degrees of freedom $(a = 5, b = 10)$;
\begin{align*}
f(x) = \frac{1}{B(a,b)} \left(\frac{a}{b}\right)^{a/2} x^{a/2 - 1} \left(1 + \frac{a}{b} x \right)^{-(a + b)/2},
\quad x > 0.
\end{align*}

Find the points $x$ such that $f(x) = 0.1$.
``` {r}
h <- 0.1
df1 <- 5
df2 <- 10
f <- function(x, h, df1, df2) { df(x, df1, df2) - h }

curve(df(x, df1, df2), xlim = c(0,6))
abline(c(0.1,0), lty = 2)

lo <- uniroot(f, interval = c(0, 1), df1 = df1, df2 = df2, h = h)
hi <- uniroot(f, interval = c(1, 3), df1 = df1, df2 = df2, h = h)
print(lo)
print(hi)
pf(hi$root, df1=df1, df2=df2) - pf(lo$root, df1=df1, df2=df2)
```

# Solving Systems of Equations
Solve the following system of equations for $x$;

\begin{align*}
&A x = b, \quad \text{where} \\
&A =
\begin{pmatrix}
4 & 1 & 0 & 0 \\
1 & 4 & 1 & 0 \\
0 & 1 & 4 & 1 \\
0 & 0 & 1 & 4
\end{pmatrix}
\quad
x =
\begin{pmatrix}
x_1 \\
x_2 \\
x_3 \\
x_4
\end{pmatrix}
\quad
b =
\begin{pmatrix}
1 \\
1 \\
1 \\
1
\end{pmatrix}
\end{align*}

``` {r}
A <- matrix(0, 4, 4)
diag(A) <- 4
diag(A[-1, -4]) <- 1
diag(A[-4, -1]) <- 1
print(A)

b <- rep(1, 4)

print(x <- solve(A, b))
A %*% x
```
