---
title: "Graphics with ggplot2"
output:
  pdf_document:
    number_sections: true
geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

Consider the `mpg` dataset.
``` {r}
tail(mpg)
```

Make an empty plot with the data prepared.
``` {r}
g <- ggplot(mpg, aes(cty, hwy))
print(g)
```

Actually plot some points.
``` {r}
g <- ggplot(mpg, aes(cty, hwy)) +
    geom_point()
print(g)
```

Add a loess smoother.
``` {r}
g <- ggplot(mpg, aes(cty, hwy)) +
    geom_point() +
    geom_smooth()
print(g)
```

Customize axes, theme, etc.
``` {r}
g <- ggplot(mpg, aes(cty, hwy)) +
    geom_point(shape = 17, size = 2) +
    geom_smooth() +
    labs(x = "City", y = "Highway", title = "Miles Per Gallon") +
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
print(g)
```

Now plot a histogram.
``` {r}
g <- ggplot(mpg, aes(hwy)) +
    geom_histogram()
print(g)
```

Customize the histogram.
``` {r}
g <- ggplot(mpg, aes(hwy)) +
    geom_histogram(bins=20, fill="blue", color="black", size=0.25)
print(g)
```

Plot a histogram for each value of `cyl`.
``` {r}
g <- ggplot(mpg, aes(hwy)) +
    geom_histogram(bins=20, fill="blue", color="black", size=0.25) +
    facet_wrap(~ cyl)
print(g)
```

Plot boxplots for each value of `class`.
``` {r}
g <- ggplot(mpg, aes(class, hwy)) +
    geom_boxplot()
print(g)
```

Plot two boxplots together horizontally.
``` {r}
library(gridExtra)

g <- ggplot(mpg, aes(class, hwy)) +
    geom_boxplot()
h <- ggplot(mpg, aes(class, cty)) +
    geom_boxplot()
grid.arrange(g, h)
```

Recall the AR(2) time series model
\begin{align*}
y_t = \phi_1 y_{t-1} + \phi_2 y_{t-2} + \epsilon_t, \quad
\epsilon_t \stackrel{\text{iid}}{\sim} \text{N}(0, \sigma^2),
\quad t = 1, \ldots, n.
\end{align*}

Generate an AR(2) series and plot it.
``` {r}
y <- arima.sim(n = 200, list(ar = c(0.5, -0.2), sd = sqrt(0.25)))
dat <- data.frame(t = 1:200, y = as.numeric(y))
g <- ggplot(dat, aes(t, y)) + geom_line()
print(g)
```

What about plotting multiple series on one plot? First draw the series.
``` {r}
n <- 200
y1 <- 0 + arima.sim(n = n, list(ar = c(0.5, -0.2), sd = sqrt(0.25)))
y2 <- 3 + arima.sim(n = n, list(ar = c(0.1, -0.2), sd = sqrt(0.25)))
y3 <- -3 + arima.sim(n = n, list(ar = c(0.7, -0.2), sd = sqrt(0.5)))
```

Make the series columns of a `data.frame`.
``` {r}
dat <- data.frame(t = 1:n, y1 = as.numeric(y1), y2 = as.numeric(y2), y3 = as.numeric(y3))
head(dat, 3)
```

Reshape the `data.frame` by stacking the series vertically.
``` {r}
library(reshape2)
newdat <- melt(dat, 't')
head(newdat)
tail(newdat)
```

Here is one way to plot the series together.
``` {r}
g <- ggplot(newdat, aes(x = t, y = value,
            group = variable,
            color = variable,
            linetype = variable)) +
    geom_line() +
    theme(legend.position = "bottom") +
    ggtitle("Multiple Series")
print(g)
```

Draw from bivariate normal.
``` {r}
library(mvtnorm)
Sigma <- matrix(c(1, 1/2, 1/2, 1), 2, 2)
x <- rmvnorm(n = 10000, mean = c(0,0), sigma = Sigma)
dat <- data.frame(x)
colnames(dat) <- c("x", "y")
```

Plot the points and superimpose contours.
``` {r}
g <- ggplot(dat, aes(x=x, y=y)) +
    geom_point() +
    geom_density2d()
print(g)
```

Plot bins instead to display density values.
``` {r}
g <- ggplot(dat, aes(x=x, y=y)) +
    geom_bin2d() +
    scale_fill_gradient(low = "blue", high = "red")
print(g)
```

Plot hexigonal bins instead.
``` {r}
library(hexbin)
g <- ggplot(dat, aes(x=x, y=y)) +
    geom_hex() +
    scale_fill_gradient(low = "blue", high = "red") +
    coord_cartesian(xlim = c(-2.5, 2.5), ylim = c(-2.5, 2.5)) +
    theme(legend.position = "bottom")
print(g)
```

Generate data from a trinomial distribution.
``` {r}
m <- 10
grid <- expand.grid(x = 0:m, y = 0:m)
grid <- grid[grid$x + grid$y <= m,]
grid$z <- m - grid$x - grid$y
```

Plot the trinomial data on a 2-d grid (no need to display the redundant third coordinate).
``` {r}
grid$dens <- apply(grid, 1, dmultinom, size=m, prob=c(0.1, 0.3, 0.6))
g <- ggplot(grid, aes(x=x, y=y)) +
    geom_raster(aes(fill = dens)) +
    scale_fill_gradient(low = "yellow", high = "red") +
    scale_x_discrete(limits = 0:m) +
    scale_y_discrete(limits = 0:m)
print(g)
```

To save the last plot, use `ggsave`. File type is determined by extension (pdf, png, jpg, etc).
``` {r, eval = FALSE}
ggsave("plot.pdf", width = 5, height = 5)
```
