---
title: "Traditional Graphics"
output: pdf_document
geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `plot` function can be used to make 2-d plots of (discrete) data.
``` {r}
x <- seq(0,1,0.1)
y <- x^2
plot(x, y)
```

We can also connect the points with lines.
``` {r}
x <- seq(0,1,0.1)
y <- x^2
plot(x, y, type = "b")
```

We can plot a function, say the normal density function `dnorm`.
``` {r}
curve(dnorm, -4, 4)
```

Create a histogram using the included `Nile` dataset.
``` {r}
print(Nile)
hist(Nile)
```

Boxplots can be created very quickly. Consider the `chickwts` dataset.
``` {r}
head(chickwts)
boxplot(weight ~ feed, data = chickwts)
```

Overlay a curve on a plot.
``` {r}
mu.hat <- mean(Nile)
sigma.hat <- sd(Nile)
hist(Nile, freq = FALSE)
curve(dnorm(x, mean = mu.hat, sd = sigma.hat), add = TRUE)
```

Instead of having the plot display in a window, request it to be written to a file using functions like `pdf`, `jpeg`, and `bmp`.
```{r, eval=FALSE}
pdf("plot.pdf", height=4, width=4)
hist(Nile)
dev.off()
```

* The function `dev.off()` closes the graphics file after writing to it.
* The units of `height` and `weight` are inches.
* To save a plot during Rstudio during interactive use, see the `Export` menu above the plot.


There are lower-level plotting functions available for drawing shapes. `abline` can be used to draw lines, e.g. to plot a simple linear regression.
``` {r}
head(cars)
fit <- lm(cars$dist ~ cars$speed)
print(fit)
plot(cars)
abline(coef = fit$coefficients)
```

Draw polygons by connecting points.
``` {r}
x <- c(1, 0, -1, 0)
y <- c(0, 1, 0, -1)
plot(x,y)
polygon(x,y)
```

Draw rectangles with `rect` and line segments with `segments`.
``` {r}
plot.new()  # Create an empty plot window, and add the parts manually
plot.window(xlim=c(-2,2), ylim=c(-2,2))
axis(1)
axis(2)
title(main="My title")
title(xlab="x-label")
title(ylab="y-label")
box()
rect(xleft = -1, ybottom = -1, xright = 1, ytop = 1, col="yellow", border = "red")
segments(x0 = -2, y0 = -2, x1 = 0, y1 = 2, col="green")
segments(x0 = 0, y0 = 2, x1 = 2, y1 = -2, col="green")
segments(x0 = -2, y0 = -2, x1 = 2, y1 = -2, col="green")
```

Multiple plots in one figure.
``` {r}
par(mfrow = c(1,2))
curve(dnorm(x), xlim = c(-3,3))
curve(pnorm(x), xlim = c(-3,3))
dev.off()
```
