---
title: "Basic programming"
author: "nagaraj neerchal"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Types 

## Arithmetic 

All arithmetic constants are fortunately available in R. 
We can 

* add 
* subtract 
* multiply 
* divide 
* exponentiate

etc

```{r}
2 + 3
2 - 3
2 * 3
pi / 2
2^3

```

Also, available are modulo arithmetic:

```{r}
21 %% 10
```

## Buit-in functions

R has many built-in functions:

```{r}
exp(1)
log(10)
log2(10)
log(10,2)
abs(-1 * 1)
sqrt(5)
```

Can generate a sequence of integers as follows:

```{r}
1:9
seq(1, 9)
seq(from = 1, to = 9)
seq(from = 1, to = 9, by = 2)
```

If function name is used without parentheses, `R` will display its definition.

```{r}
seq
```


The result of an expression can be assigned it to a named variable with the `<-` or `=` operators.

```{r}
x <- 2
y <- x + 3
y
x
x = 2
y = x + 4
y
x
```

..or increment x as follows:

```{r}
x <- x + 1
x
```

...or more complicated expressions like $y = y+x^2+cos(x)+1/x$ as follows:

```{r}
y <- y+x^2+cos(x)+1/x
y
```

## Vectors

Three basic functions for constructing vectors are

```{r}
print(x <- seq(1, 20, by = 2))
print(y <- rep(3, 4))
print(z <- c(y, x))
```
Notice we used the `print()` function to avoid repeating the object on a separate line for the default print method.

Some people are switching away from the `<-` assignment operator to a simple `=`. The rarely used ambiguity for assignment in function parameters is addressed with parentheses.
```{r}
print(x = seq(5, 20, by = 5))
x
print((x = seq(5, 20, by = 5)))
x
```

The : operator takes precedence over operators such as `+` and `-`. Use parentheses when in doubt, or to improve readability.

```{r}
n <- 10
print(x <- 1:n + 1)
print(x <- 1:(n + 1))
```

Indexing into vectors

```{r}
x <- seq(10, 20)
x[9]
x[12]        # Past the end
x[2:5]
ind <- c(1, 3, 5)
x[ind]
x[-1]
x[-c(1, 3, 5, 7)]
x[x > 11]
ind <- which(x > 11)
ind
x[ind]
```

Arithmetic on Vectors

```{r}
x <- c(1, 2, 3)
y <- c(4, 5, 6)
x + y            # Most operations are elementwise.
y - x
x * y
y / x
y^x
1:4 + 1:2        # Unequal length: shorter vector is repeated.
1 + 1:10         # Scalar plus vector
c(1, 2, 3) + c(1, 2)
```

Some Useful Functions on Vectors

```{r}
x <- c(5, 4, 3, 6, 7, 8, 9, 10, 1, 2)
sort(x)
x <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
rev(x)
x <- seq(1, 10)
mean(x)
sum(x)
prod(x)
y <- sample(x)
min(y)

# Draw a sample (with or without replacement) from x at random
x <- seq(1, 10)
sample(x)    
sample(x, size = 5)
sample(x, replace = TRUE)
sample(x, replace = TRUE, size = 5)
```

## Other Data Types:

```{r}
# Complex numbers

a <- complex(5, 3); b <- complex(7, 2)
a + b

# Logicals

x <- c(TRUE, TRUE, FALSE); y <- c(TRUE, FALSE, FALSE)
x & y
!x
c(sum(x), prod(x))

#Strings

a <- "my string"
length(a)
nchar(a)
x <- 5
cat("The answer is", x, ".\n")

str <- sprintf("The answer is %d.", x)
print(str)
```


Factors represent categorical data. They give descriptive labels, and also emphasize that values are not numbers.


```{r}
x <- sample(1:3, size = 10, replace = TRUE)
x
y <- factor(x, labels=c("urban", "suburban", "rural"))
y
levels(y)


#Splitting numerical data into intervals, using the \code{cut} function.
x <- rnorm(10)
x
cut(x, 3)
```

Some scalar values with special meanings:


```{r}
y <- log(0)
y
is.infinite(y)
is.na(y)
x <- log(-1)
is.na(x)
z <- c(1, 2, 3, NA, 4)
is.na(z)
mean(z)
mean(z, na.rm = TRUE)
```

To avoid issues due to rounding error, use `all.equal` to test for numerical equality.

```{r}
2 * 2 == 4
sqrt(2) * sqrt(2) == 2
all.equal(sqrt(2) * sqrt(2), 2)
```


## Matrics

```{r}

# Creating a 3 by 3 matrix of NA's 

matrix(NA, nrow = 3, ncol = 3)

# Creating a 2 by 2 matrices from a given vector of data

matrix(c(1, 2, 3, 4), nrow = 2, ncol = 2)
matrix(c(1, 2, 3, 4), nrow = 2, ncol = 2, byrow = TRUE)

# A diagonal matrix with custom labels

A <- diag(5, 3)        
rownames(A) <- c("First Row", "Second Row", "Third Row")
colnames(A) <- c("A", "B", "C")
print(A)
nrow(A)
ncol(A)
```

Indexing Matrices


```{r}
# Access some specific elements

A <- matrix(seq(1, 9), nrow = 3, ncol = 3)
A
A[2, 3]
A[3, 2]

# First row

A[1, ] 

# Second column

A[,2]        

# Rows 1 and 2

A[c(1, 2), ]   

# Columns 2 and 3 of row 1

A[1, c(2, 3)] 
```

Matrix Operations

```{r}

A <- B <- matrix(seq(1, 9), nrow = 3, ncol = 3)

# Addition 

A + B

# Multiply by a scalar

3 * A        

# Multiply two matrices

A            
B
A %*% B

A <- matrix(c(1, 2, 3), nrow = 3, ncol = 3)
A

# Transpose

t(A)                
A <- matrix(c(1, 0, 0, 0, 2, 0, 0, 0, 3), nrow = 3, ncol = 3)
A

# Inverse of non-singular matrix

solve(A)          

# Kronecker product

x <- matrix(c(1, 2, 3), 1, 3)
y <- matrix(c(1, 0, 0), 1, 3)
x %x% y             


# Eigen value/vector decomposition
# Singular valud decomposition

A <- matrix(c(1, 0, 0, 0, 2, 0, 0, 0, 3), nrow = 3, ncol = 3)
eigen(A)    
svd(A)        
```


Stacking Matrices


```{r}
A <- matrix(0, nrow = 3, ncol = 3)
B <- matrix(1, nrow = 3, ncol = 3)

rbind(A, B)
cbind(A, B)
```


Functions to operate on matrices:

```{r}

# sqrt works elementwise on matrix

A <- matrix(seq(1, 9), nrow = 3, ncol = 3)
sqrt(A)            

# Note: see the chol function to compute B such that B %*% B = A

# sum and max treat the matrix like a vector

A <- matrix(seq(1, 9), nrow = 3, ncol = 3)
A
sum(A)            
max(A)

# Apply max to each row of A

apply(A, 1, max)

# Apply to columns of A

apply(A, 2, max)
```

## Arrays

Arrays are similar to matrices, but can have $>2$ dimensions.

```{r}
A <- array(seq(1, 27), dim = c(3, 3, 3))
A

# 2nd col from 1st slice

A[, 2, 1] 

# 1st slice

A[, , 1]   

# 1st and 2nd slice

A[, , 1:2]

# Apply the function sum to 3rd slice

apply(A, 3, sum)
```

## Lists

* Vectors, matrices, and arrays are intended for scalars, usually numbers.
* Lists can hold elements of varying types.
* Lists can be used to construct more complicated data structures.


```{r}
x <- list(a = "string value", b = 5.5, c = TRUE)
print(x)

x$a

x[["a"]]

x[[1]]
```

Lists are used to return multiple results from a function See e.g. `eigne`


```{r}
eigen
```


Data frames are convenient for data in table format. Each column represents a variable, and can have its own data type.

The `CO2` dataset comes with `R`. First, check the help page using `?CO2`.

```{r}
#?CO2

# Now let us view the dataset

CO2

# Access each field as a vector using `$` operator.

CO2$uptake

# Access rows or columns using matrix indexing.
 
CO2[3, ]
CO2[, 4]
CO2[3, 4]


# Create a new `data.frame`


data.frame(new.conc = CO2$conc, new.uptake = CO2$uptake)

```

Working with dataframes


```{r}
# Sorting a dataframe by a column

idx <- order(CO2$uptake)
CO2[idx,]

# Merging with ``inner join'' of data.frames, like SQL.

id <- 1:4
color <- c("red", "green", "blue", "yellow")
data1 <- data.frame(id=id,color=color)
data1

x <- 2:5
y <- c("square", "trianlge", "circle", "star")
z <- c("a", "b", "c", "d")
data2 <- data.frame(id = x, shape = y, letter = z)
data2

merge(data1, data2, by = "id")

# For more examples of manipulating data.frames, see www.statmethods.net/management

```


## Functions

```{r}
# Functions are objects and can be passed like any other data type in R.

f <- function(x) { x + 1 }
f(1)
g <- function(f, x) { f(x)^2 }
g(f, 1)

# We can also explicitly specify what to return

f <- function(x) {
  if (x %% 2 == 0)
    return("even")
  else
    return("odd")
}

# You can run the above function as follows

# source("function.R")
print(f(1))
print(f(2))

# Can set default values for arguments.
f <- function(x, y, z = 100) { c(x, y, z) }
f(y = 1, x = 3)

# Argument types can be checked manually.
f <- function(x) { stopifnot(is.numeric(x)); return(x^2) }
f(1)

# The above code is designed to terminate when user supplies an illegal argument
# Hence the code will terminate and hence building of the Rmarkdown file. 
# tryCatch function is used to prevent the exit

tryCatch(f("a"),
    error = function(e) {
        return(e)
    }
)


# Functions can be constructed around variables that are defined outside. This can be a source of confusion.

y <- 5
f <- function(x) { x + y }
f(1)
```


## Control Flow


```{r}
# if statement

u <- runif(1)
if (u < 0.25) {
    y <- 0
} else if (u < 0.5) {
    y <- 1
} else {
    y <- 2
}
# switch statement

x <- "1"
switch(x,
    "1" = "blue",
    "2" = "green",
    "3" = "red",
    "default")

# for loop

for (i in 1:10) { cat("Iteration", i, "\n") }

# while loop

u <- 1
while (u > 0.1) {
    u <- runif(1)
}

# try/catch statement


f <- function(x) { stop("Error!!") }
ret <- tryCatch(f(5),
    error = function(e) {
        return(e)
    }
)

## ret is the message e

g <- function(x) { x }
ret <- tryCatch(g(5),
    error = function(e) {
        return(e)
    }
)

## ret is 5




#  apply statement


X <- matrix(rnorm(1000*5), 1000, 5)
apply(X, 2, sd)


# sapply statement


f <- function(x) { if (x %% 2 == 0) { "even" } else { "odd" } }
x <- rpois(8, 5)
sapply(x, f)

sapply(x, f)



# \code{lapply} statement

L <- list()
L[[1]] <- rnorm(100)
L[[2]] <- rnorm(100)
L[[3]] <- rnorm(100)

lapply(L, function(x) { c(length(x), mean(x), sd(x)) })
```








